
\begin{abstract}
The {\em Cold Dead Fish} system (CDF), also known as the {\em Coherent
Description System}, seeks to extend logic programming with facilities
for ontology management.  CDF includes features standard to
frame-based ontology systems, including support for inheritance,
updates, and a visual editor.  CDF furthermore checks consistency of
ontologies by means of both logic programming and theorem-proving
techniques, incorporating advantages of ${\cal ALC}$-like description
logics.  Features novel to CDF are its integration of
description-logic style specifications with Prolog-style querying, a
module-like component system with support for concurrent editing of
ontologies, and a flexible rule interface that supports, among other
features, a lazy database access mechanisms.  CDF has been used to
construct many commercial ontology-based systems for textual
extraction and understanding, and many of its features are
open-source.  The semantics of CDF ontologies is formalized along with
the relation between CDF ontologies and description logics.  Finally,
results on preservation of submodels in Type-0 ontologies are
presented, providing a starting point for formalizing ontology
updates.
\end{abstract}

\mycomment{
In order to meet these dual goals, CDF distinguishes between Type 0
and Type 1 ontologies.  Type 1 ontologies represent classes using an
expressive language in which common problems are $P$-space complete.
However many ontologies do not require the full expressive power of
Type 1 ontologies, so Type 0 ontologies are introduced as a
restriction in which common reasoning problems require only polynomial
time.  This paper presents a first-order logical semantics for Type 0
ontologies, and characterizes when such ontologies are consistent and
when they have free models (in which equality does not hold for any
two distinct elements of the universe).  Type 0 ontologies contain
features that are not common in description logics, such as product
identifiers, non-inheritable relations and individuals.  These
features are formalized, and the relation between Type 0 ontologies
and description logics is shown.  Finally, results on preservation of
submodels in Type-0 ontologies are presented, providing a starting
point for formalizing ontology updates.

Cdf, the {\em Cold Dead Fish System}, is an experimental platform to
combine features of description logics and frame-based ontology
systems with logic programming.
}