\section{Configuring and Examining the CDF System} \label{sec:config} 

The file \file{cdf\_config.P} contains all facts, tables etc. that may
need to be configured for a particular user or application.  In
addition to the predicates described below, {\tt cdf\_config.P}
contains a fact for the dynamic predicate {\tt
default\_user\_error\_handler/1} which is used to handle errors.  See
the XSB manual for documentation of this predicate.

\subsection{The Configuration API} \label{sec:configapi}

\begin{description}

%--------------------------------------------------------------------
\mycomment{
\ourpredmoditem{cdf\_flag/2}{usermod}
{\tt cdf\_flag(?Flag,?Value} is a dynamic predicate that indicates all
aspects of a CDF state that can be set dynamically at run time.  It is
analogous to the XSB predicate {\tt xsb\_flag/2}.  The aspects that
can be dynamically changed include consistency checks to be set for
various contexts.

\ourpredmoditem{initial\_cdf\_flag/2}{usermod}
The main action of {\tt initial\_cdf\_flag(+Flag,+ActionList)} is used
to declare which actions (usually consistency checks) are to be
executed at specific points during ontology management.  Any {\tt
initial\_cdf\_flag/2} fact is meant for site-wide configuration.
It can be overridden by declaring a corresponding {\tt
user\_cdf\_flag/2} fact in {\tt \$HOME/.xsb/xsbrc.P}.  
}
%--------------------------------------------------------------------

\ourpredmoditem{cdf\_configuration/2}{usermod}
{\tt cdf\_configuration(?Flag,?Value)} allows access to configuration
parameters for CDF, and is analogous to {\tt xsb\_configuration/2}
These parameters are set during the compilation of CDF by define
statements in \file{cdf\_definitions.h}.  Currently, the only
parameters that can be set are
\begin{itemize}
\item {\tt using\_xj}.  This parameter is set {\tt on} if XJ is to be
supported by CDF, and set {\t off} otherwise.  This parameter is set on by
default in the proprietary version of CDF and off by default in the
open-source version.

\item {\tt tabled\_isa}.  This parameter is set on if {\tt isa/2} is
tabled, and off otherwise.  It is set on by default.
\end{itemize}

\ourpredmoditem{cdf\_index/3}{usermod}
{\tt cdf\_index(+Functor,+Arity,+Index)} is used to set initial
indices for the various types of extensional facts. These indices can
be changed, if necessary to give better performance, but note that the
semantics of predicates in the Type-0 API may not use all indices.
See \refsec{sec:type0api} for details.

\ourpredmoditem{initialize\_state/0}{usermod}
Normally, initialization is done automatically upon loading CDF at the
start of a session.  The predicate {\tt initialize\_state/0} should be
called only when a state is to be reinitialized during a session.
This predicate removes all data in extensional and intensional forat,
and reasserts the basic CDF classes and relations, and resets internal
state variables to values in the CDF configuration file.

\end{description}

