\chapter{Implementation and System Features} \label{sec:impl}

In this section we describe how the CDF system implements the
semantincs of Part~\ref{part:semantics}, as well as many other
features for ontology management.  We begin by describing CDF
identifiers, facts, and rules in Section~\ref{sec:instance}.  The
Type-0 query interface, based on tabled resolution is described in
Section~\ref{sec:type0query}.  Section~\ref{sec:type0query} describes
the Type-1 API, which is based on the CDF therorem prover.  Update and
consistency predicates for CDF are described in
Section~\ref{sec:update}.  Section~\ref{sec:config} describes how to
configure CDF, along with predicates that allow the user to examine
aspects of the CDF state.  Section~\ref{sec:components} describes
basic I/O for CDF, along with a more sophisticated {\em component}
system that is built on top of basic I/O.  Section~\ref{sec:database}
describes database interfaces for CDF, and
Section~\ref{sec:concurrency} describes concurrency support.
%
\mycomment{
   Finally Section~\ref{} briefly describes the InterProlog interface for
   XSB, the XJ interface that is built using InterProlog, the CDF caching
   system that supports visual editing of CDF instances, and the CDF
   Editor.}

\paragraph{Terminology}
Throughout this chapter, we assume a general familiarity with the
terminology and conventions of Prolog in general and XSB in
particular.  We discuss here our specialized conventions.  By
\emindex{sorts} we mean the various sorts for identifiers discussed in
\refchap{chap:type0} and in \refdef{def:cdfids}.  We reserve the term
\emindex{types} for types as defined in the ISO-Prolog standard
\cite{ISO}.  We use \emindex{domains} to refer to the ``types'' that are
useful in CDF.  Each CDF domain is defined by a 1-ary predicate that
begins with {\tt is}, such as {\tt isContextType/1}, {\tt
isType0Term/1} etc., and we refer to domains via their predicate names
and arities (e.g. {\tt isContextType/1}, etc).


%----------------------------------------------------
\input{instances} 
%-------------------------------------------
\input{type0impl}
%------------------------------------------------------------------
\input{type1impl}
%------------------------------------------------------------------
\input{update}
%-----------------------------------------------------------------------------
\input{checks}
%-----------------------------------------------------------------------------
\input{configuration}
%------------------------------------------------------------------------------
\input{components}
%------------------------------------------------------------------------------
\input{database}
%------------------------------------------------------------------------------
\input{concurrency}
